An Empirical Analysis of the Role of Energy in Economic Growth

### 2014-04-24 Article/EconGrowth2.Rnw
The article now uses the modified historicalPlot function.

<<setup, echo=FALSE, message=FALSE, eval=TRUE>>=
require(EconModels)
require(lattice)
require(latticeExtra)
require(ggplot2)
require(car)
require(mosaic)
require(xtable)
require(nlmrt)
require(micEconCES)
require(plyr)
# Provides access to data melting. 
# See http://cran.r-project.org/web/packages/reshape2/reshape2.pdf
require(reshape2) 
# tikz allows use of LaTeX formatting and font in graphs. 
# Allows for a consistent look across the paper.
# See http://r-forge.r-project.org/R/?group_id=440 for instructions 
# on installing tikzDevice.
# require(tikzDevice) 
trellis.par.set(theme=theme.mosaic())
options(width=75)
options(show.signif.stars=TRUE)
require(knitr)
opts_chunk$set(
  # ********************************************************
  eval=TRUE,
  # ********************************************************
  # tikz allows LaTeX code in graphical output. 
  # E.g., "$y$" for a variable name in the legend for a graph.  
  # dev='tikz',
  tidy=FALSE,
  comment=NA,
  # Tells whether to cache output from chunks are saved. Cacheing saves time. 
  # However, references to LaTeX tables DO NOT WORK if cacheing is turned on.
  cache=FALSE,
  warning=FALSE,  #Tells whether to show warnings in the output.
  # Turns off messages for all chunks. 
  # Set TRUE on an individual chunk to see it.
  message=FALSE, 
  # Tells whether to echo code for all chunks. 
  # Set TRUE on an individual chunk to see its code.
  echo=FALSE     
)
@

<<constants, include=FALSE>>=
# Graph Heights
# #' @export
maxHeight <- 8.5 # Inches
# #' @export
ninePanelScatterGraphHeight <- maxHeight # Inches
# #' @export
# ninePanelTernaryGraphHeight <- 7 # Inches
# #' @export
# ninePanelSpaghettiGraphHeight <- 7 # Inches

# Graph Widths
# #' @export
maxWidth <- 6.5 # Inches
# #' @export
fourPanelScatterGraphWidth <- maxWidth # Inches
# #' @export
# threePanelScatterGraphWidth <- maxWidth*0.75 # Inches
# #' @export
# onePanelScatterGraphWidth <- maxWidth*0.25
# #' @export
# fourPanelTernaryGraphWidth <- maxWidth
# #' @export
# twoPanelTernaryGraphWidth <- maxWidth*0.45
# #' @export
# fourPanelSpaghettiGraphWidth <- maxWidth
# #' @export
# twoPanelSpaghettiGraphWidth <- maxWidth*0.55
# #' @export
# onePanelSpaghettiGraphWidth <- maxWidth*0.4

#' @export
presentationGraph1RowHeight <- 2.8 # Inches
#' @export
presentationGraph2ColWidth <- 2 # Inches
#' @export
presentationGraph1ColWidth <- 2*presentationGraph2ColWidth
# #' @export
# keyTextSize <- 0.85 # 85% of normal size
# #' @export
# keyColumns <- 1 # Want only 1 column in the key for lattice graphs
# #' @export
# defaultKeyXLoc <- 0.01 # x position of the key. This default is good for 9-panel graphs.
# #' @export
# defaultKeyYLoc <- 0.95 # y position of the key. This default is good for 9-panel graphs.
# Parameter graphs are tricky to ensure same size. I've found that with height set to 5 in for the 
# cd graphs and 2.93 in for the sf graphs, I get a nearly-exact match with graph size. From these settings I 
# can set up 2 equations:
#    5 in = 2*hgraph + haxis
#    2.93 in = hgraph + haxis
# to estimate that h_axis = 0.86 in and hgraph = 2.05 in.
# #' @export
h_param_graph <- 2.07 # Inches for a single panel
# #' @export
h_param_axis <- 0.86 # Inches (also includes whatever margin is applied)
# #' @export
# sfParameterGraphHeight <- h_param_graph + h_param_axis
# #' @export
cdParameterGraphHeight <- 2*h_param_graph + h_param_axis
# #' @export
cesParameterGraphHeight <- 3*h_param_graph + h_param_axis
# #' @export
linexParameterGraphHeight <- h_param_graph + h_param_axis
# #' @export
sfParameterGraphWidth <- maxWidth
# #' @export
cdParameterGraphWidth <- maxWidth
# #' @export
# cdTernaryGraphWidth <- maxWidth/2.2
# #' @export
# cdSpaghettiGraphWidth <- maxWidth/2.2
# #' @export
cesParameterGraphWidth <- maxWidth
# #' @export
linexParameterGraphWidth <- maxWidth
# Other graph parameters that apply to all graphs
# #' @export
scaleTextSize <- 0.8  # Multiple of normal size
# #' @export
scaleTickSize <- -0.5 # 50% of normal size and pointing INWARD!

spaghettiGraphLevel <- 0.95 # Spaghetti graph bands will encompass 95% of the data.

# In Rstudio, this paper is generated from it's directory, which is "Article".
# We need to define paths relative to the location of the paper. 
# Relative path to historical data from Article
baseHistorical <- file.path("..", "data")
# We also need to define a directory path to the resample data results from "Article"
baseResample <- file.path("..", "data_resample")
@

@@

\author[CalvinEngr]{Caleb Reese}
\author[CalvinEngr]{Lucas Timmer}
\author[CalvinEngr]{Matthew Kuperus Heun\corref{cor1}}
\ead{mkh2@calvin.edu}
\author[Stellenbosch]{Martin de Wit}
\author[CalvinMath]{Randall J. Pruim}

@@

\twocolumn{}
<<Historical_Lattice_Graph, fig.pos="H", fig.align='center', fig.width=3, fig.height=7.5, fig.cap="Historical data: indexed GDP ($y$), capital stock ($k$), labor ($l$), and thermal energy ($q$).">>=
historicalPlot(baseHistorical=baseHistorical)
@

@@

Figure~\ref{fig:Historical_Lattice_Graph} presents historical data, 
including indexed values for 
output ($y$), 
capital stock ($k$), 
labor ($l$), and 
energy ($q$) 
for 1980--2011 (developed economies) and 1991--2011 (other economies). 
Factors of production track each other closely, 
and high covariances among these variables can be seen in 
Tables~\ref{tab:Covariance_US}--\ref{tab:Covariance_ZM}. 
\authNote{Do we want to discuss the covariances?}

>>>>

<<setup, echo=FALSE, message=FALSE, eval=TRUE>>=
require(EconModels2)
require(lattice)
require(latticeExtra)
require(ggplot2)
require(car)
require(mosaic)
require(xtable)
require(nlmrt)
require(micEconCES)
require(plyr)
# Provides access to data melting. 
# See http://cran.r-project.org/web/packages/reshape2/reshape2.pdf
require(reshape2) 
# tikz allows use of LaTeX formatting and font in graphs. 
# Allows for a consistent look across the paper.
# See http://r-forge.r-project.org/R/?group_id=440 for instructions 
# on installing tikzDevice.
# require(tikzDevice) 
trellis.par.set(theme=theme.mosaic())
options(width=75)
options(show.signif.stars=TRUE)
require(knitr)

# Pull in some constants and other important functions
source("Scripts/files.R")

# ********************************************************
# We need to define a directory path to the resample data results from "Article"
archive <- file.path("..", "data_resample_2014-04-24_Refactored-Code_n=1000.zip")
baseResample <- "data_resample" # Define the directory within the archive wherein we can find the data.
# ********************************************************
# Load the fits to the original historical data out of the archive we're using.
# This object should be passed in to functions that need specific fits.
f <- gzcon(unz(archive, file.path(baseResample, "oModels.Rdata")))
oModels <- readRDS(f)
close(f)
# In Rstudio, this paper is generated from it's directory, which is "Article".
# We need to define paths relative to the location of the paper. 
# Relative path to historical data from Article
baseHistorical <- file.path("..", "data")

opts_chunk$set(
  # ********************************************************
  eval=TRUE,
  # ********************************************************
  # tikz allows LaTeX code in graphical output. 
  # E.g., "$y$" for a variable name in the legend for a graph.  
  # dev='tikz',
  tidy=FALSE,
  comment=NA,
  # Tells whether to cache output from chunks are saved. Cacheing saves time. 
  # However, references to LaTeX tables DO NOT WORK if cacheing is turned on.
  cache=FALSE,
  warning=FALSE,  #Tells whether to show warnings in the output.
  # Turns off messages for all chunks. 
  # Set TRUE on an individual chunk to see it.
  message=FALSE, 
  # Tells whether to echo code for all chunks. 
  # Set TRUE on an individual chunk to see its code.
  echo=FALSE     
)
@

<<constants, include=FALSE>>=
source(file.path("..", "Scripts", "Constants.R"))

# Graph Heights
maxHeight <- 8.5 # Inches
ninePanelScatterGraphHeight <- maxHeight # Inches
# ninePanelTernaryGraphHeight <- 7 # Inches
# ninePanelSpaghettiGraphHeight <- 7 # Inches

# Graph Widths
maxWidth <- 6.5 # Inches
fourPanelScatterGraphWidth <- maxWidth # Inches
# threePanelScatterGraphWidth <- maxWidth*0.75 # Inches
# onePanelScatterGraphWidth <- maxWidth*0.25
# fourPanelTernaryGraphWidth <- maxWidth
# twoPanelTernaryGraphWidth <- maxWidth*0.45
# fourPanelSpaghettiGraphWidth <- maxWidth
# twoPanelSpaghettiGraphWidth <- maxWidth*0.55
# onePanelSpaghettiGraphWidth <- maxWidth*0.4

presentationGraph1RowHeight <- 2.8 # Inches
presentationGraph2ColWidth <- 2 # Inches
presentationGraph1ColWidth <- 2*presentationGraph2ColWidth
# keyTextSize <- 0.85 # 85% of normal size
# keyColumns <- 1 # Want only 1 column in the key for lattice graphs
# defaultKeyXLoc <- 0.01 # x position of the key. This default is good for 9-panel graphs.
# defaultKeyYLoc <- 0.95 # y position of the key. This default is good for 9-panel graphs.
# Parameter graphs are tricky to ensure same size. I've found that with height set to 5 in for the 
# cd graphs and 2.93 in for the sf graphs, I get a nearly-exact match with graph size. From these settings I 
# can set up 2 equations:
#    5 in = 2*hgraph + haxis
#    2.93 in = hgraph + haxis
# to estimate that h_axis = 0.86 in and hgraph = 2.05 in.
h_param_graph <- 2.07 # Inches for a single panel
h_param_axis <- 0.86 # Inches (also includes whatever margin is applied)
# sfParameterGraphHeight <- h_param_graph + h_param_axis
cdParameterGraphHeight <- 2*h_param_graph + h_param_axis
cesParameterGraphHeight <- 3*h_param_graph + h_param_axis
linexParameterGraphHeight <- h_param_graph + h_param_axis
sfParameterGraphWidth <- maxWidth
cdParameterGraphWidth <- maxWidth
cesParameterGraphWidth <- maxWidth
linexParameterGraphWidth <- maxWidth
# Other graph parameters that apply to all graphs
scaleTextSize <- 0.8  # Multiple of normal size
scaleTickSize <- -0.5 # 50% of normal size and pointing INWARD!

spaghettiGraphLevel <- 0.95 # Spaghetti graph bands will encompass 95% of the data.
@

@@

\author[CalvinEngr]{Matthew Kuperus Heun\corref{cor1}}
\ead{mkh2@calvin.edu}
\author[Stellenbosch]{Martin de Wit}
\author[CalvinMath]{Randall J. Pruim}
\author[CalvinEngr]{Caleb Reese}
\author[CalvinEngr]{Lucas Timmer}

@@

\twocolumn{}
<<Historical_Graph, fig.pos="H", fig.align='center', fig.width=3, fig.height=7.5, fig.cap="Historical data: indexed GDP ($y$), capital stock ($k$), labor ($l$), and thermal energy ($q$).">>=
historicalData <- read.table(file=file.path(baseHistorical, "AllData.txt"), header=TRUE)
levels(historicalData$Country) <- countryAbbrevs
historicalPlot(historicalData)
@

@@

Figure~\ref{fig:Historical_Graph} presents historical data, 
including indexed values for 
output ($y$), 
capital stock ($k$), 
labor ($l$), and 
energy ($q$) 
for 1980--2011 (developed economies) and 1991--2011 (other economies). 
Factors of production track each other closely, 
and high covariances among these variables can be seen in 
Tables~\ref{tab:Covariance_US}--\ref{tab:Covariance_ZM}. 
\authNote{Do we want to discuss the covariances?}
